---
title: "STUtility - Installation"
author:
date: ''

---

<style type="text/css">
div.main-container {
  background-color: #000000 !important;
  max-width: 1400px;
  margin-left: auto;
  margin-right: auto;
}
</style>
<style>
#TOC {
  background: url("https://www.spatialresearch.org/wp-content/uploads/2019/09/str-logo-spatial_research_3@2x.png");
  background-size: contain;
  padding-top: 100px !important;
  background-repeat: no-repeat;
  op: 5%;
  opacity: 0.8;
  width: 500px;
  color: white;
  border-color: #000000 !important;
}
</style>

<style> code, pre{
  background-color: #000000 !important;
  color: white !important;
}
</style>
<style> 
body {
  color: white
}
</style>




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Installation

Using devtools, execute the following from the R console:
<br>
```{r, eval=FALSE}

devtools::install_github(
    "jbergenstrahle/STUtility"
   , build_opts = c("--no-resave-data", "--no-manual")
)

```


### Note on MacOS

If you run into `Error in dyn.load(file, DLLpath = DLLpath, ...):`, make sure you have XQuartz installed, which can be downloaded from [xquartz.org](https://www.xquartz.org/)

&nbsp;
<hr />
<p style="text-align: center;">A work by <a href="j.bergenstrahle@scilifelab.se">Joseph Bergenstråhle</a> and <a href="ludvig.larsson@scilifelab.se">Ludvig Larsson</a></p>

<!-- Add icon library -->
<link rel="stylesheet" href="https://cdnjs.cloudflare.com/ajax/libs/font-awesome/4.7.0/css/font-awesome.min.css">

<!-- Add font awesome icons -->
<p style="text-align: center;">
    <a href="https://www.spatialresearch.org" class="fa fa-beer"></a>
</p>

&nbsp;
