---
title: "Installation"
author:
date: ''

---

<style type="text/css">
div.main-container {
  background-color: #FFFFFF !important;
  max-width: 1800px;
  margin-left: auto;
  margin-right: auto;
}
</style>

<style>
#TOC {
  background: url("https://www.spatialresearch.org/wp-content/uploads/2019/09/str-logo-spatial_research_3@2x.png");
  background-size: contain;
  padding-top: 100px !important;
  background-repeat: no-repeat;
  width: 500px;
  color: black;
  border-color: #FFFFFF !important;
}
</style>

<style> code, pre{
  background-color: #F8F8F8 !important;
  color: black !important;
}
</style>

<style> .watch-out {
  background-color: #F0F0F0;
  border: 1px #F0F0F0;
}
<style>

<style> 
body {
  color: black;
  font-size: 20pt;
}
</style>




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Install

To install STutility from github (currently only option), you need to have `devtools` installed:

```{r, eval=FALSE}

install.packages("devtools")

```

Using devtools, execute the following from the R console to install STutility:

```{r, eval=FALSE}

devtools::install_github("jbergenstrahle/STUtility")

```


### Common issues on Mac OS

If you run into `Error in dyn.load(file, DLLpath = DLLpath, ...):`, make sure you have XQuartz installed, which can be downloaded from [xquartz.org](https://www.xquartz.org/)

If you are missing gfortran you can get it from [CRAN](https://cran.r-project.org/bin/macosx/tools/)


### Set up conda environment (Mac OS)

Below are instructions for how to set up a conda environment to install STutility. 

From the terminal

```{r, eval=FALSE}

conda create -n R4.0
conda activate R4.0
conda install -c conda-forge r-essentials r-base r-devtools r-spdep r-hdf5r
conda install -c bioconda r-fftwtools

```

From R

```{r, eval=FALSE}

devtools::install_github("jbergenstrahle/STUtility")

```


&nbsp;
<hr />
<p style="text-align: center;">A work by <a href="j.bergenstrahle@scilifelab.se">Joseph Bergenstråhle</a> and <a href="ludvig.larsson@scilifelab.se">Ludvig Larsson</a></p>

<!-- Add icon library -->
<link rel="stylesheet" href="https://cdnjs.cloudflare.com/ajax/libs/font-awesome/4.7.0/css/font-awesome.min.css">

<!-- Add font awesome icons -->
<p style="text-align: center;">
    <a href="https://www.spatialresearch.org" class="fa fa-beer"></a>
</p>

&nbsp;
