---
title: "Manual Image Alignment"
author:
date: ''
---

<style type="text/css">
div.main-container {
  background-color: #FFFFFF !important;
  max-width: 1800px;
  margin-left: auto;
  margin-right: auto;
}
</style>

<style>
#TOC {
  background: url("https://www.spatialresearch.org/wp-content/uploads/2019/09/str-logo-spatial_research_3@2x.png");
  background-size: contain;
  padding-top: 100px !important;
  background-repeat: no-repeat;
  width: 500px;
  color: black;
  border-color: #FFFFFF !important;
}
</style>

<style> code, pre{
  background-color: #F8F8F8 !important;
  color: black !important;
}
</style>

<style> .watch-out {
  background-color: #F0F0F0;
  border: 1px #F0F0F0;
}
<style>

<style> 
body {
  color: black;
  font-size: 20pt;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, autodep = TRUE)
```

```{r load_lib, warning=FALSE, message=FALSE}
library(STutility)
```


# Manual image alignment

```{r}
se.list <- readRDS("~/HER2ST_TMP/results/R_objects/BC.seu.list")
```


```{r}
se <- ManualAlignImages(se.list[[2]])
```

