---
title: "Neighborhood analysis"
author:
date: ''
---

<style type="text/css">
div.main-container {
  background-color: #FFFFFF !important;
  max-width: 1800px;
  margin-left: auto;
  margin-right: auto;
}
</style>

<style>
#TOC {
  background: url("https://www.spatialresearch.org/wp-content/uploads/2019/09/str-logo-spatial_research_3@2x.png");
  background-size: contain;
  padding-top: 100px !important;
  background-repeat: no-repeat;
  width: 500px;
  color: black;
  border-color: #FFFFFF !important;
}
</style>

<style> code, pre{
  background-color: #F8F8F8 !important;
  color: black !important;
}
</style>

<style> .watch-out {
  background-color: #F0F0F0;
  border: 1px #F0F0F0;
}
<style>

<style> 
body {
  color: black;
  font-size: 20pt;
}
</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, autodep = TRUE)

```

```{r load_lib, warning=FALSE, message=FALSE, include=FALSE}
library(STutility)
```


## Region neighbors
***

Sometimes it can be useful to extract the "neighborhood" of a set of spots. As an example, we show how this can be applied to find all the neighboring spots of any region of interest. 

To demonstrate the regional neighbors feature, we will use the Breast Cancer data available on the 10x Genomics [website](https://support.10xgenomics.com/spatial-gene-expression/datasets).

The Seurat object below contain 2 breast cancer tissue sections which have already been normalized and clustered. 

<br>
```{r load_BC_data, include = FALSE}

load("pre_data/preSaved_10x_BC.RData")

```
```{r plot_BC_data, fig.width=14, fig.height=6}
FeatureOverlay(se, features = "seurat_clusters", sampleids = 1:2, ncols = 2)
```
<br>

### Connected Spatial Network

Once you have defined a region of interest and you want to find all spots neighboring to this region you can use the `RegionNeighbours` function to automatically detect such spots. 

For example, let’s say that we want to select all neighbors to cluster 2. The first step is to make sure that the identity of your seurat object is correct, here we need to set it to "seurat_clusters".
<br>
```{r find_region_neighbors}

se <- SetIdent(se, value = "seurat_clusters")
se <- RegionNeighbours(se, id = "2", verbose = TRUE)

```
<br>
The default behavior is to find all spots which are neighbors with the selected id but ignoring its label, therefore it will simply be called nbs_2 as in “neighbors to 2”. The output will be stored as a new column in the meta.data slot, and in this example will be called "nbs_2". The neighborhood detection algorithm is applied to each section separately and can therefore be run on multiple sections at the same time.
<br>
```{r find_neighbors, fig.width = 14, fig.height=6, out.width = "100%"}

FeatureOverlay(se, features = "nbs_2", ncols = 2, sampleids = 1:2, cols = c("red", "lightgray"), pt.size = 2)

```
<br>
You can also keep all spots from the id group by setting keep.within.id = TRUE.
<br>
```{r region_neighbors_within}

se <- SetIdent(se, value = "seurat_clusters")
se <- RegionNeighbours(se, id = 2, keep.within.id = T, verbose = TRUE)

```

```{r find_neighbors_2, fig.width = 14, fig.height=6, out.width = "100%"}

FeatureOverlay(se, features = "nbs_2", ncols = 2, sampleids = 1:2, cols = c("red", "lightgray"), pt.size = 2)

```
<br>

Using these two sets of spots, we can run a DE analysis to check what genes are up-regulated outside the cluster border.

<br>
```{r de_neighbors, fig.width=10, fig.height=10}

library(magrittr)
library(dplyr)

se <- SetIdent(se, value = "nbs_2")
nbs_2.markers <- FindMarkers(se, ident.1 = "2", ident.2 = "nbs_2")
nbs_2.markers$gene <- rownames(nbs_2.markers)
se.subset <- SubsetSTData(se, expression = nbs_2 %in% c("2", "nbs_2"))
sorted.marks <- nbs_2.markers %>% top_n(n = 40, wt = abs(avg_logFC))
sorted.marks <- sorted.marks[order(sorted.marks$avg_logFC, decreasing = T), ]
DoHeatmap(se.subset, features = sorted.marks$gene, group.colors = c("red", "lightgray"), disp.min = -2, disp.max = 2)

```
<br>

From this DE-test we can for example see that the genes COX6C and FCGR3B genes are up-regulated inside the cluser whereas LGALS1 and CYBA genes are more highly expressed outisde the cluster border.

<br>
```{r plot_markers, fig.width = 14, fig.height=12, out.width = "100%"}

FeatureOverlay(se.subset, features = c("COX6C", "FCGR3B", "LGALS1", "CYBA"), pt.size = 2,  
               ncols = 2, cols = c("darkblue", "cyan", "yellow", "red", "darkred"))

```
<br>

And lastly, if you want to keep the labels for the neighbors you can set keep.idents = TRUE and you can keep one label per identity for the neighboring spots, e.g. "label"_nb_to_2

<br>

```{r, echo=T, fig.width = 14, fig.height=6, out.width = "100%"}

se <- SetIdent(se, value = "seurat_clusters")
se <- RegionNeighbours(se, id = 2, keep.idents = TRUE, verbose = TRUE)
FeatureOverlay(se, features = "nbs_2", ncols = 2, sampleids = 1:2, pt.size = 2)

```

&nbsp;
<hr />
<p style="text-align: center;">A work by <a href="mailto: j.bergenstrahle@scilifelab.se">Joseph Bergenstråhle</a> and <a href="mailto: ludvig.larsson@scilifelab.se">Ludvig Larsson</a></p>

<!-- Add icon library -->
<link rel="stylesheet" href="https://cdnjs.cloudflare.com/ajax/libs/font-awesome/4.7.0/css/font-awesome.min.css">

<!-- Add font awesome icons -->
<p style="text-align: center;">
    <a href="https://www.spatialresearch.org" class="fa fa-beer"></a>
</p>

&nbsp;
