---
title: "Plots and themes"
author:
date: ''
---

<style type="text/css">
div.main-container {
  background-color: #FFFFFF !important;
  max-width: 1800px;
  margin-left: auto;
  margin-right: auto;
}
</style>

<style>
#TOC {
  background: url("https://www.spatialresearch.org/wp-content/uploads/2019/09/str-logo-spatial_research_3@2x.png");
  background-size: contain;
  padding-top: 100px !important;
  background-repeat: no-repeat;
  width: 500px;
  color: black;
  border-color: #FFFFFF !important;
}
</style>

<style> code, pre{
  background-color: #F8F8F8 !important;
  color: black !important;
}
</style>

<style> .watch-out {
  background-color: #F0F0F0;
  border: 1px #F0F0F0;
}
<style>

<style> 
body {
  color: black;
  font-size: 20pt;
}
</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, autodep = TRUE)

```

```{r load_lib, warning=FALSE, message=FALSE, include=FALSE}
library(STutility)
```
```{r load_lib_hidden, include=FALSE}
library(kableExtra)
library(magrittr)
```

```{r, pre-load, echo=FALSE, eval=TRUE, include=FALSE}

load("pre_data/preSaved_10xHippo_norm_reductions.RData")

```

## Patchwork
***

Many of the plots generated with vsiualization functions from STutility are built using the [patchwork](https://github.com/thomasp85/patchwork) R package. This package makes it much easier to change the layout and themes of different plots and we'll go through a couple of examples here.

Let's draw a spatial distribution of Pvalb and Th using `ST.FeaturePlot` and a violin plot showing the expression of these genes within each clusters.

Some of the layout options can be controlled direclty from `ST.FeaturePlot` using for example `ncol` and `grid.ncol`, but you can also rearrange the plot afterwards. Here we set `ncol = 2` to specify that the sections will be arranged in two columns and `grid.ncol = 1` to specify that the features will be arranged in 1 column.

<br>
```{r plot1, fig.width=10, fig.height=10}

p1 <- ST.FeaturePlot(se, features = c("Pvalb", "Th"), ncol = 2, grid.ncol = 1)

p1

```

Now let's add a violin plot and show it side by side with the spatial feature plot.

```{r plot2, fig.width=12, fig.height=8}

p2 <- VlnPlot(se, features = c("Pvalb", "Th"), ncol = 1, group.by = "seurat_clusters")

p1 - p2

```

As you can see, it is very easy to combine plots side by side. If you want the sub plots to take up more or less area of the total plot, you can specify layout options with the patchwork function `plot_layout`.


```{r plot3, fig.width=12, fig.height=8}

p1 <- ST.FeaturePlot(se, features = c("Pvalb", "Th"), grid.ncol = 1, indices = 1)
p2 <- VlnPlot(se, features = c("Pvalb", "Th"), ncol = 1, group.by = "seurat_clusters")

# Give the second plot with a width that is 2x the width of the first
p1 - p2 + patchwork::plot_layout(widths = c(1, 2))

```

Or an even more complex example

```{r plot4, fig.width=14, fig.height=8}

p3 <- ST.FeaturePlot(se, features = c("Pvalb", "Th"), ncol = 2, grid.ncol = 2, show.sb = FALSE)
p1 <- FeaturePlot(se, features = c("Pvalb", "Th"), cols = c("mistyrose", "red", "darkred"))
p2 <- VlnPlot(se, features = c("Pvalb", "Th"), ncol = 2, group.by = "seurat_clusters")

(p1 - p2)/p3

```

## Themes
***

It is also easy to change the theme of your plots, even after it has been drawn. You can specify a custom theme using the `custom.theme` argument in `ST.FeaturePlot`, `FeatureOverlay`, etc. But it's even easier with the patchwork system.

```{r plot5, fig.width=10, fig.height=5}

custom_theme <- theme(legend.position = c(0.45, 0.8), # Move color legend to top
                      legend.direction = "horizontal", # Flip legend
                      legend.text = element_text(angle = 30, hjust = 1), # rotate legend axis text
                      strip.text = element_blank(), # remove strip text
                      plot.title = element_blank(), # remove plot title
                      plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "cm")) # remove plot margins

p <- ST.FeaturePlot(se, features = "nFeature_RNA", ncol = 2, show.sb = FALSE, palette = "Spectral")
p & custom_theme

```

Or you can for example add a grid to show the x/y axes. Here, the x/y axes reoresent the pixel coordinates mapped to the "tissue_hires_image.png" from the spaceranger output.

```{r plot6, fig.width=11, fig.height=5}
p & theme_bw()
```

<br>


&nbsp;
<hr />
<p style="text-align: center;">A work by <a href="mailto: j.bergenstrahle@scilifelab.se">Joseph Bergenstråhle</a> and <a href="mailto: ludvig.larsson@scilifelab.se">Ludvig Larsson</a></p>

<link rel="stylesheet" href="https://cdnjs.cloudflare.com/ajax/libs/font-awesome/4.7.0/css/font-awesome.min.css">

<p style="text-align: center;">
    <a href="https://www.spatialresearch.org" class="fa fa-beer"></a>
</p>

&nbsp;
